---
title: "Rountine analysis"
output: rmarkdown::html_vignette
description: >
  the example rountine analysis 
vignette: >
  %\VignetteIndexEntry{Rountine analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# <font size="5">Metabolite compound name transformation</font>

## <font size="4">Metabolite compound name transform to RefMet name</font>

<font color="CadetBlue">**This step requires networking**</font>

**RefMet: A Reference list of Metabolite names.**The main objective of RefMet is to provide a standardized reference nomenclature for both discrete metabolite structures and metabolite species identified by spectroscopic techniques in metabolomics experiments.

```{r message=FALSE, warning=FALSE}
library(MNet)
library(dplyr)
library(tibble)
library(survival)
library(tidyr)
library(knitr)
library(stringr)
library(ggplot2)
library(RColorBrewer)
library(clusterProfiler)
library(org.Hs.eg.db)
library(pathview)
```

```{r message=FALSE, warning=FALSE}
compound_name <- c("2-Hydroxybutyric acid",
                   "1-Methyladenosine",
                   "tt",
                   "2-Aminooctanoic acid")
## transform the compound name to refmet name
refmetid_result <- name2refmet(compound_name)
head(refmetid_result)
```

## <font size="4">Metabolite compound name transform to KEGG ID</font>

<font color="CadetBlue">**This step requires networking**</font>

Transform the metabolites compound name to KEGG ID

```{r message=FALSE, warning=FALSE}
compound_name <- c("2-Hydroxybutyric acid",
                   "1-Methyladenosine",
                   "tt",
                   "2-Aminooctanoic acid")
## transform the compound name to KEGG ID, some metabolites have several KEGG ID
keggid_result <- name2keggid(compound_name) %>%
  separate_rows(KEGG_id, sep = ";")
head(keggid_result)
```

## <font size="4">Metabolite name corresponding to kegg pathway</font>

<font color="CadetBlue">**This step requires networking**</font>

Search the kegg pathway corresponding to the metabolite name

```{r message=FALSE, warning=FALSE}
compound_name <- c("2-Hydroxybutyric acid",
                   "1-Methyladenosine",
                   "tt",
                   "2-Aminooctanoic acid")
## Search the kegg pathway corresponding to the metabolite name
result_all <- name2pathway(compound_name)
head(result_all)

##### Output is the each metabolite related pathway
result_name2pathway <- result_all$name2pathway
head(result_name2pathway)

## the KEGG ID of the metabolite name
result_name2keggid <- result_all$kegg_id
head(result_name2keggid)

## the pathway of the metabolite name
result_name2enrichpathway <- result_all$pathway
head(result_name2enrichpathway)
```

## <font size="4">Metabolite KEGG ID transform to KEGG pathway</font>

KEGG ID transform to KEGG pathway

```{r message=FALSE, warning=FALSE}
keggid <- c("C05984", "C02494")
##### the output is the each metabolite related pathway
keggpathway_result <- keggid2pathway(keggid)
head(keggpathway_result)
```

## Pathway information
Get the gene and the metabolite in the pathway

```{r message=FALSE, warning=FALSE}
## the genes and metabolites in pathway 'hsa00630'
result <- pathwayinfo("hsa00630")

## the genes and metabolites in pathway 'Glyoxylate and dicarboxylate metabolism'
result <- pathwayinfo("Glyoxylate and dicarboxylate metabolism")

head(result$gene_info[1:2, ])
head(result$compound_info[1:2, ])
```

## <font size="4">Pathway name transform to pathway id</font>
Transform the KEGG pathway name to KEGG pathway ID

```{r message=FALSE, warning=FALSE}
## the KEGG pathway ID of pathway name
pathwayid <- pathway2pathwayid("Glycolysis / Gluconeogenesis")
head(pathwayid)
```

# <font size="5">Group-wise analyses</font>
## <font size="4">PCA plot</font>

PCA of the data

```{r fig.keep='all', PCA, fig.width=10.00, fig.height=6.18, message=FALSE, warning=FALSE}
### the pca plot
p_PCA <- pPCA(meta_dat, group)
p_PCA$p1
p_PCA$p2
p_PCA$p3
```

## <font size="4">Differnetial metabolite analysis</font>

### <font size="3">Function 'mlimma'</font>
```{r message=FALSE, warning=FALSE}
## mlimma is the function of Differential Metabolite analysis by limma
diff_result <- mlimma(meta_dat, group)
head(diff_result)
```

### <font size="3">Function 'DM'</font>

```{r message=FALSE, warning=FALSE}
## DM is the function of Differential Metabolite analysis by OPLS-DA
diff_result <- DM(2 ** meta_dat, group)
head(diff_result)

## filter the differential metabolites by default fold change >1.3 or < 1/1.3 ,fdr < 0.05 and VIP>0.8

diff_result_filter <- diff_result %>%
  filter(Fold_change > 1.3 | Fold_change < 1 / 1.3) %>%
  filter(Padj_wilcox < 0.1) %>%
  filter(VIP > 0.8)
head(diff_result_filter)
```


## <font size="4">Differential metabolites' volcano</font>

Volcano plot of metabolites using the function "pVolcano"

```{r fig.keep='all', pVolcano, fig.width=10.00, fig.height=6.18, message=FALSE, warning=FALSE}
p_volcano <- pVolcano(diff_result, foldchange_threshold = 1.5)
p_volcano
```

## <font size="4">Differential metabolites' heatmap</font>

Heatmap plot of differentital metabolites using the function "pHeatmap"

```{r fig.keep='all', pHeatmap, fig.width=10.00, fig.height=6.18, message=FALSE, warning=FALSE}
meta_dat_diff <- meta_dat[rownames(meta_dat) %in% diff_result_filter$Name, ]
p_heatmap <- pHeatmap(
  meta_dat_diff,
  group,
  fontsize_row = 5,
  fontsize_col = 4,
  clustering_method = "ward.D",
  clustering_distance_cols = "correlation"
)
```

## <font size="4">Differential metabolites' zscore</font>

Zscore plot of differentital metabolites using the function "pZscore"

```{r fig.keep='all', pZscore, fig.width=10.00, fig.height=6.18, message=FALSE, warning=FALSE}
p_zscore <- pZscore(meta_dat_diff, group, ysize = 3)
p_zscore
```

# <font size="5">Feature selection</font>

## <font size="4">Boruta</font>
Using machine learning "Boruta" for feature selection 

```{r message=FALSE, warning=FALSE}
#group <- rep("normal",length(names(meta_dat)))
#group[grep("TUMOR",names(meta_dat))] <- "tumor"
meta_dat1 <- t(meta_dat) %>%
  as.data.frame() %>%
  mutate(group = group)

result_ML_Boruta <- ML_Boruta(meta_dat1)
head(result_ML_Boruta)
```

## <font size="4">Random Forest</font>
Using machine learning "Random Forest" for feature selection

```{r message=FALSE, warning=FALSE}
result_ML_RF <- ML_RF(meta_dat1)
result_ML_RF$p
result_ML_RF$feature_result
```

## <font size="4">XGBoost</font>
Using machine learning "  XGBoost" for feature selection

```{r message=FALSE, warning=FALSE}
result_ML_xgboost <- ML_xgboost(meta_dat1)
result_ML_xgboost$p
result_ML_xgboost$feature_result
```

## <font size="4">LASSO</font>
Using machine learning "LASSO" for feature selection

```{r message=FALSE, warning=FALSE}
result_ML_lasso <- ML_alpha(meta_dat1, method = "lasso")
head(result_ML_lasso)
```

## <font size="4">elastic network</font>
Using machine learning "elastic network" for feature selection

```{r message=FALSE, warning=FALSE}
result_ML_elastic <- ML_alpha(meta_dat1, method = "elastic")
head(result_ML_elastic)
```
# <font size="5">Pathway analysis</font>

## <font size="4">Pathview only metabolites</font>

```{r eval=FALSE, message=FALSE, warning=FALSE}
kegg_id <- c("C02494", "C03665", "C01546", "C05984", "C14088", "C00587")
value <- c(-0.3824620,
           0.1823628,
           -1.1681486,
           0.5164899,
           1.6449798,
           -0.7340652)
names(value) <- kegg_id
cpd.data <- value

gene_name <- c("LDHA", "BCKDHB", "PCCA", "ACSS1")
gene_value <- c(1, 0.5, -1, -1)
names(gene_value) <- gene_name

## pathview plot of metabolites
pPathview(cpd.data, outdir = "result_v0131")
```

## <font size="4">Pathview contains metabolites and genes</font>

```{r eval=FALSE, message=FALSE, warning=FALSE}
## pathview plot of genes and metabolites
pPathview(cpd.data = cpd.data,
          gene.data = gene_value,
          outdir = "result_v0131")
```

# <font size="5">Clinical analysis</font>
## <font size="4">Time series of clinical</font>
Column contains the time, group, clinical index(such as ALT), low and high
```{r fig.keep='all', pCliTS, fig.width=10.00, fig.height=6.18, message=FALSE, warning=FALSE}
clinical_index[1:5, ]
time_series_ALT <- pCliTS(clinical_index, "ALT")
time_series_ALT
```

## <font size="4">Survival analysis</font>
```{r fig.keep='all', survCli, fig.width=10.00, fig.height=6.18, message=FALSE, warning=FALSE}
names(aml)[3] = "group"
p = survCli(aml)
p
```

## <font size="4">Metabolites’ survival plot</font>

```{r eval=FALSE, message=FALSE, warning=FALSE}
metabolites <- c("C03819", "C02918", "C03916")
survMet(dat_surv,
        metabolites,
        cluster_method = "mean",
        out_dir = "result_v0131/survival/")
```

## <font size="4">Cox analysis</font>

```{r message=FALSE, warning=FALSE}
result <- MetCox(dat_surv)
head(result)
```

# <font size="5">Session information</font>

```{r message=FALSE, warning=FALSE}
sessionInfo()
```
