--- 
title: "MNet"
author: "Tuantuan Gui, Xuan Liu and Hai Fang"
date: "`r Sys.Date()`"
description: >
  This is a minimal example of using the bookdown package to write a book.
  The HTML output format for this example is bookdown::gitbook,
  set in the _output.yml file.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MNet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## <font size="5">R Installation</font>

R (http://www.r-project.org) is a language and environment for statistical computing and graphics. We assume R (version 4.0.0 or higher) has been installed in your local machine. The latest version can be installed following instructions below for different platforms (Windows, Mac, and Linux).

- Quick link for Windows: [Download R for Windows](https://cran.r-project.org/bin/windows/base/).

- Quick link for Mac: [Download R for Mac OS X 11](https://cran.r-project.org/bin/macosx/big-sur-arm64/base/).

- Below are shell command lines in Terminal (for Linux):

```{bash,eval=FALSE}
#Ensure you have the conda:
conda install -c conda-forge r-base
```

## <font size="5">MNet Installation</font>

MNet requires R version 4.0.0 or higher, please see https://cran.r-project.org.

MNet is available for all operating systems and can be installed via the Github.

```{r,eval=FALSE}
if(!require(BiocManager)){
  install.packages("BiocManager")
}

if (!require(remotes)) {
  BiocManager::install("remotes", dependencies=T)
}

if (!require(devtools)) {
  BiocManager::install("devtools", dependencies=T)
}

BiocManager::install("hfang-bristol/dnet", dependencies=T)
BiocManager::install("tuantuangui/MNet", dependencies=T)

#If the installation of MNet is failed, please install the failed dependency package again, such as if the ERROR is "ERROR: dependencies ‘xxx’ is not available for package ‘MNet’", please install the 'xxx' package again.
```

Check the package 'MNet' successfully installed
```{r,eval=FALSE}
library(help=MNet)
```
## <font size="5">Frequently Asked Questions</font>
1. Can not install dependent packages dnet  
If the ERROR is "Error: Failed to install 'dnet' from GitHub: Could not resolve host: api.github.com", please try it again.

```{r,eval=FALSE}
BiocManager::install("hfang-bristol/dnet", dependencies=T)
```

## <font size="5">Usage</font>

- Extended pathway analysis can be found [here](https://tuantuangui.github.io/MNet/articles/01-Pathway_analysis.html)

- Metabolism-related subnetwork analysis can be found [here](https://tuantuangui.github.io/MNet/articles/02-Subnetwork_analysis.html)

- Routine analysis can be found [here](https://tuantuangui.github.io/MNet/articles/03-Routine_analysis.html)
